In Class Work: Plotting Conditional Means
----------------------------

Today we're going to work again with attrition as our outcome. Your company would like to know which groups are most likely to leave this year. Your supervisor has asked you to wtrie a one-page memo, detailing how attrition varies as a function of different employee characteristics. Since your supervisor is a visual learner, the memo should include two graphics, with accompanying brief text. No R code should be visible in the memo, just graphics and normal text. Your supervisor has said that the following areas are of key interest as possible predictors of attrition.

1. Commute distance
2. Marital Status
3. Gender
4. Work/Life Balance
5. Education level
6. Salary


```{r  include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(forcats)
library(RColorBrewer)
```

```{r}
at<-read_csv("https://community.watsonanalytics.com/wp-content/uploads/2015/03/WA_Fn-UseC_-HR-Employee-Attrition.csv")

## Save for later
save(at,file="at.Rdata")

## Load the saved version (not necessary, just showing we can!)
load("at.Rdata")
View(at)
```

```{r}
#so that I hide everything except graphics and actual text
echo=FALSE 
results='hide' 
warning=FALSE 
message=FALSE
```

```{r}
## Crate a new variable named attrit and define it as 0
at<-at%>%mutate(attrit=ifelse(Attrition=="Yes",1,0))

table(at$Attrition)

table(at$attrit)

table(at$attrit,at$Attrition)

```

```{r}

at_sum<-at%>%
  group_by(DistanceFromHome)%>%
  summarize(attr_avg=mean(attrit))

at_sum

```

```{r distance from home}
## Bar Plot with aesthetics: mean attrition as height, distance from home as cateogry
gg<-ggplot(at_sum,aes(x=fct_reorder(as.factor(DistanceFromHome),attr_avg),y=attr_avg))
## Use bar plot geometry and height of bars set by level observed in dataset
gg<-gg+geom_bar(stat="Identity")
#label
gg<-gg+xlab("Distance From Home")+ylab("Yearly Attrition")
## Print
gg
```

```{r marital status}
at_sum<-at%>%
  group_by(MaritalStatus)%>%
  summarize(attr_avg=mean(attrit))

at_sum

## ordered by level
gg<-ggplot(at_sum,aes(x=fct_reorder(MaritalStatus,attr_avg),y=attr_avg))

gg<-gg+geom_bar(stat="identity")

## Labeling
gg<-gg+xlab("Marital Status")+ylab("Yearly Attrition")
##Print
gg
```

```{r gender}
at_sum<-at%>%
  group_by(Gender)%>%
  summarize(attr_avg=mean(attrit))

at_sum

## ordered by level
gg<-ggplot(at_sum,aes(x=fct_reorder(Gender,attr_avg),y=attr_avg))

gg<-gg+geom_bar(stat="identity")

## Labeling
gg<-gg+xlab("Gender")+ylab("Yearly Attrition")
##Print
gg
```
```{r work life balance}
at_sum<-at%>%
  group_by(WorkLifeBalance)%>%
  summarize(attr_avg=mean(attrit))

at_sum

## ordered by level
gg<-ggplot(at_sum,aes(x=fct_reorder(as.factor(WorkLifeBalance),attr_avg),y=attr_avg))

gg<-gg+geom_bar(stat="identity")

## Labeling
gg<-gg+xlab("Work Life Balance")+ylab("Yearly Attrition")
##Print
gg
```
```{r education level}
at_sum<-at%>%
  group_by(Education)%>%
  summarize(attr_avg=mean(attrit))

at_sum

## ordered by level
gg<-ggplot(at_sum,aes(x=fct_reorder(as.factor(Education),attr_avg),y=attr_avg))

gg<-gg+geom_bar(stat="identity")

## Labeling
gg<-gg+xlab("Education Level")+ylab("Yearly Attrition")
##Print
gg
```
```{r salary}
at_sum<-at%>%
  group_by(MonthlyIncome)%>%
  summarize(attr_avg=mean(attrit))

at_sum

## ordered by level
gg<-ggplot(at_sum,aes(x=fct_reorder(MonthlyIncome,attr_avg),y=attr_avg))

gg<-gg+geom_bar(stat="identity")

## Labeling
gg<-gg+xlab("Salary")+ylab("Yearly Attrition")
##Print
gg
```

```{r}
## Summarize attrition by travel AND gender
at_sum<-at%>%
  group_by(BusinessTravel,Gender)%>%
  summarize(attr_avg=mean(attrit))

## Get the results
at_sum

## Plot it using a bar plot
gg<-ggplot(at_sum,aes(x=reorder(BusinessTravel,attr_avg),y=attr_avg,color=Gender))
gg<-gg+geom_bar(stat="identity",aes(fill=Gender),position="dodge")
gg