---
output:
  word_document: default
  pdf_document: default
  html_document: default
---
#Presenting Data 1: Plotting Conditional Means {#plot_means}

The idea when plotting conditional means is to show how the outcome, or variable of interest, varies as a function of predictors. 

Today we'll be working with a dataset from IBM which provide a standard HR dataset, which we can use to predict attrition. Attrition in this case is defined as an employee leaving without being fired or retiring. Companies generally attempt to avoid attrition, as it's very expensive to search for and hire a replacement-- better in general to keep the employees you have, provided they are doing their jobs. This means that it's important to predict who might leave in a given year. This information can be used in a targeted way in order to focus resources on the employees most likely to leave. 

## Setup for plotting conditional means

We start with a standard set of setup commands. Today we'll be working with `tidyverse`, as usual, along with a library called `forcats` which helps us to deal with the dreaded factor variables. To handle colors, we'll need the package `RColorBrewer.`

```{r  include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(forcats)
library(RColorBrewer)
```

Next we load in the data, using the `readr` package. Note that this data is saved as comma separated or `csv.` This is an easy file format to recognize. When we use the `readr` package, it gives us some output that says how it interprets the data-- is it a string variable, numeric (float), integer and so on. 
-csv means comma separated version because it separates different elements with a comma
-what you save your data as is arbitrary cuz it's only on your computuer, but you should make it the same name as what it's loaded as so that you're not confused

## Loading Data 
```{r}
at<-read_csv("https://community.watsonanalytics.com/wp-content/uploads/2015/03/WA_Fn-UseC_-HR-Employee-Attrition.csv")

## Save for later
save(at,file="at.Rdata")

## Load the saved version (not necessary, just showing we can!)
load("at.Rdata")
View(at)
```

Today, our primary outcome of interest will be attrition. This is a binary variable that is currently encoded as text-- "Yes" or "No." We need to encode it as a binary variable with 1 meaning yes and 0 meaning no. After recoding, we need to make sure that the new variable looks correct.
-should always use a binary variable where 1 means yes and 0 means no; if you were going to do something like gender you'd do "female" and 1 would be yes and 0 would be male (as opposed to saying "gender" and then the 1 and 0 are confusing)

```{r}
## Crate a new variable named attrit and define it as 0
at<-at%>%mutate(attrit=ifelse(Attrition=="Yes",1,0))

table(at$Attrition)

table(at$attrit)

table(at$attrit,at$Attrition)

```

In this recoding, we use "if else" to give a logical statement. In this case, we're saying if the file "attrition" is a yes, then give it a 1. The comma 0 means if it doesn't say yes (meaning it says no), then give it a zero
-needed to start with at and then the overwrite of at so that R knows which data set we're working with
-we made a new variable "attrit" so that we could keep the "Attrition" variable as yes and no

##Univariate Graphics

Univariate graphics help us understand what individual variables look like-- how are they distibuted across the sample? Here's a quick rundown on some univariate graphics. Say we wanted a quick count of who was in each department. We can use geom_bar to get this done. By default, this will give us a count in each department. 
-when you say "fill" and then give the =department, that means that you're going to fill the chart based on department
-the colors are assigned automatically, but we can change those later
-the initial command of using the "at" dataset is imperative so that gg knows where we're pulling data
-the second line of code (69) says use gg with the geom bar (which automatically gives count as the y axis)
-never use pie charts (they don't really convey accurate information) or 3D charts (they add unnecessary dimensions to your data) or anything wonky like that

```{r}
gg<-ggplot(at,aes(x=Department,fill=Department))
gg<-gg+geom_bar()
gg
``` 

The next univariate graphic you should know is for continuous variables. The first thing you generally want is a histogram. 

```{r}
gg<-ggplot(at,aes(x=DistanceFromHome))
gg<-gg+geom_histogram(binwidth = 1,fill="lightblue")
gg
``` 
This histogram shows us that most employees live super close to work (less than 5 miles away even)
-this means it's "right skewed data" and the mean is higher


Density plots provide a continous graphic of the distribution of a variable:
```{r}
gg<-ggplot(at,aes(x=DistanceFromHome))
gg<-gg+geom_density()
gg

## Changing bandwidth-- not recommended, just showing you how. 
gg<-gg+geom_density(bw=.2)
gg
```
This density plot shows the same info as the histogram; it's showing how many people fall into each range 
If you want to follow the data more closely, lower the increment of your bandwidth. The reason it looks bizarre in the one with .2 bandwidth is because the data wasn't reported in increments less than a mile, so the data spikes at each increment that is a decimal of the number (like the high point is the people who travel 5 miles and the trough is no one because no one reported 5.4 miles etc)


## Predicting Attrition
Our first prediction will use business travel as a predictor for attrition. There are three categories here-- non travel, travel infrequently, and frequent travel. We'll calculate levels of attrtion at teach level and then take a look at the data. 

```{r}

at_sum<-at%>%
  group_by(BusinessTravel)%>%
  summarize(attr_avg=mean(attrit))

at_sum

```

Remember that the mean of a binary variable indicates the proportion of the population that has a certain characteristcs. So, in our case, `r  round(at_sum$attr_avg[2],2)` of the sample that travels frequently left the company in the last year. Our first plot will be a basic bar plot, showing the average levels of attrition. 

```{r}
## Bar Plot with aesthetics: mean attrition as height, business travel as cateogry
gg<-ggplot(at_sum,aes(x=BusinessTravel,y=attr_avg))
## Use bar plot geometry, height of bars set by level observed in dataset
gg<-gg+geom_bar(stat="Identity")
## Print
gg
```

This is fine, but it should really be in the order of the underlying variable. We can use `fct_reorder` to do this. 

_Side_ _Note_

What is a factor variable? In R, factor variables are used for categorical data. These are data elements that can take on one and only one value of a mutually exclusive and exhaustive list of elements. In our case, the travel variable is a factor-- employees can be in Non-Travel, Travel Frequently or Travel Rarely bins. Everyone is one bin, and the bins cover all possible options. We use factors when numbers won't work-- for characteristics like race or religion or political affiliation. 

```{r}
## Same asethetics, but now orderred by level
gg<-ggplot(at_sum,aes(x=fct_reorder(BusinessTravel,attr_avg),y=attr_avg))

gg<-gg+geom_bar(stat="identity")

## Labeling
gg<-gg+xlab("Amount of Travel")+ylab("Yearly Attrition")
##Print
gg
```
that first line of code says reorder the x axis based on business travel in order of the attr_avg
-always order the barcharts in order of the underlying variable; it's much easier to understand (esp as we get more complicated)
-if you wanted to reorder from highest to lowest, put a - in front of the attr_avg in the code


*Quick Exercise: Create a bar plot showing average attrition by department instead of travel*
There's two steps to this: find the averages of each bin using attrition (called at_sum (with attrit as a binary variable)) and group it by Department, then plot Department (ordered by attrition rate average) with attrit on the y axis
```{r}

at_sum<-at%>%
  group_by(Department)%>%
  summarize(attr_avg=mean(attrit))

at_sum

```

```{r}
## Same asethetics, but now orderred by level
gg<-ggplot(at_sum,aes(x=fct_reorder(Department,attr_avg),y=attr_avg))

gg<-gg+geom_bar(stat="identity")

## Labeling
gg<-gg+xlab("Department")+ylab("Yearly Attrition")
##Print
gg
```

A dot plot can be a good way of displaying conditional means as well. Many times dot plots are more easily understood if they are horizontal, so we'll use `coord_flip` to make it horizontal.

```{r}

at_sum<-at%>%
  group_by(BusinessTravel)%>%
  summarize(attr_avg=mean(attrit))

at_sum
## Now a dot plot
gg<-ggplot(at_sum,aes(x=reorder(BusinessTravel,attr_avg),y=attr_avg))
gg<-gg+geom_point()
gg<-gg+coord_flip()
gg
```

*Quick Exercise: Create a dot plot showing average attrition by department*

## Conditional means using two predictors

We can use graphics to display conditonal means at multiple levels of predictor levels. There are a couple of ways to get this done. When using bar plots we've got two basic tools: location and color. In the first example, we're going to plot attrition by travel and gender, We'll use color to indicate gender, and location to indicate travel.


```{r}
## Summarize attrition by travel AND gender
at_sum<-at%>%
  group_by(BusinessTravel,Gender)%>%
  summarize(attr_avg=mean(attrit))

## Get the results
at_sum

## Plot it using a bar plot
gg<-ggplot(at_sum,aes(x=reorder(BusinessTravel,attr_avg),y=attr_avg,color=Gender))
gg<-gg+geom_bar(stat="identity",aes(fill=Gender),position="dodge")
gg

## Plot it using a dot plot
gg<-ggplot(at_sum,aes(x=reorder(BusinessTravel,attr_avg),y=attr_avg),color=Gender)
gg<-gg+geom_point(aes(color=Gender))
gg<-gg+coord_flip()
gg


```
The "dodge" position means set the bars next to each other instead of splitting the bars on top of each other

*Quick Exercise: Create either a bar plot or a dot plot showing attrition by department AND field of education*

## MOAR Variables: faceting

We can continue this logic with three variables. Now we're going to summarize by Travel, Gender and Marital status. Here we're going to use an additional tool in our arsenal: Faceting. Faceting means making multiple graphs with the same structure. In the code below, we will arrange positions based on travel, color based on gender, and then split the graphic by marital status. 

```{r}

at_sum<-at%>%
  group_by(BusinessTravel,Gender,MaritalStatus)%>%
  summarize(attr_avg=mean(attrit))

at_sum

at_sum%>%arrange(attr_avg)

gg<-ggplot(at_sum,aes(x=reorder(BusinessTravel,attr_avg),
                      y=attr_avg,
                      fill=Gender))
## Bar plot, with unstacked (dodge)
 gg<-gg+geom_bar(stat="identity",position="dodge")
## Separate out by Marital Status 
gg<-gg+facet_wrap(~MaritalStatus)
## Change orientation to sideways
gg<-gg+coord_flip()
## Print
gg
```
The facet_wrap code means "create that same graphic but arrange it by this characteristic." In that way, we created different columns based on marital status in which you can still see travel and attrition
-just use the tilda ~ with the characteristic you want to group by
-the stat identity code means plot the actual numbers (it's your aesthetic)

_Side_ _Note_: Colors

What if we want to change colors? This is a little tricky for most people at first. `ggplot` thinks in terms of palettes, so you need to associate a palette with a characteristics of the graphic. Below, I replace the default palette with my own ugly one. 

```{r}

## Changing Colors
mypal<-c("lightblue","orange")

gg<-gg+scale_fill_manual(values =mypal )
## Print
gg
```
R already has a large palette of colors so it'll recognize what you say in quotes, but not all colors will work like that
-the scale_fill_manual command is how you use the palette you just created (this time we named our palette "mypal," but it could have been anything)

You can also use `RColorBrewer` which has a wide variety of palettes already built. Below I use the qualitative palette creatively named "Set1".
-these are pre-buile palettes, as opposed to creating our own palette like we did above
```{r}
## Another way, using color brewer palettes: 
gg<-gg+scale_fill_brewer(palette = "Set1")

gg

```

*Quick Exercise: Plot predicted attrition by Education Field, Department and Gender*

## Multiple Predictors for Conditional Means

Once you get past three variables, things can get difficult. One solution is to create a new factor with one level for every single level of the predictor variables. 

```{r ,fig.height=6}
##This gets a little nutty
at_sum<-at%>%
  group_by(BusinessTravel,Gender,MaritalStatus,WorkLifeBalance)%>%
  summarize(attr_avg=mean(attrit))%>%
  ungroup()%>%
  arrange(attr_avg)

at_sum

#this table we just created is way too big, so we need another way to group
## One Solution: a new variable for every type of employee defined by travel, gender, marital status, and happiness with work/life

at_sum$grouping<-paste0(at_sum$BusinessTravel,
                       ", ",
                       at_sum$Gender,
                       ", ",
                       at_sum$MaritalStatus,
                       ", Work/Life:",
                       at_sum$WorkLifeBalance)
#line 293 means that we take the dataset at_sum and select the variable "grouping," which we just created and defined as the four variables we want
#paste0 combines text separated by a comma (or space depending on default option)
at_sum$grouping<-as.factor(at_sum$grouping)

at_sum%>%select(grouping,attr_avg)

at_sum<-at_sum%>%filter(attr_avg>.01)

gg<-ggplot(at_sum,aes(x=fct_reorder(grouping,attr_avg),y=attr_avg))
gg<-gg+geom_bar(stat="identity",aes(fill=MaritalStatus))
gg<-gg+coord_flip()
gg

##Cleaning up a bit

gg<-ggplot(at_sum,aes(x=fct_reorder(grouping,attr_avg),y=attr_avg))
gg<-gg+geom_bar(stat="identity",aes(fill=MaritalStatus))
gg<-gg+ylab("Proportion of Employees Who Departed")+xlab("Category")
gg<-gg+coord_flip()
gg
```

The other solution is to use facets, or lots of little graphs, which show how the pattern varies across different groups. In this case, our groups will be defined by gender and work/life balance. 

```{r}
## Using Facets

at_sum<-at%>%
  group_by(BusinessTravel,Gender,MaritalStatus,WorkLifeBalance)%>%
  summarize(attr_avg=mean(attrit))%>%
  ungroup()%>%
  arrange(attr_avg)

at_sum


gg<-ggplot(at_sum,aes(x=fct_reorder(BusinessTravel,attr_avg),y=attr_avg))
gg<-gg+geom_bar(stat="identity",aes(fill=MaritalStatus),position="dodge")
gg<-gg+facet_wrap(~Gender+WorkLifeBalance,ncol=4)


gg<-gg+ylab("Proportion of Employees Who Departed")+xlab("Category")
gg<-gg+theme(axis.text.x = element_text(angle = 60, hjust = 1))
gg<-gg+ggtitle("Departure by Gender and Level of Work/Life Satisfaction")
gg

```
The issue is that both the macro barchart and the combo of multiple charts are fairly difficult to read and explain, so it's best to not use this many variables unless you have a compelling reason or a really clear relationship of the data that will show on a graphic

*Sort of Quick Exercise: Try and Replicate one of the above plots using performance review, department, education field and overtime. *
