---
title: "05-regression_inclass_cook_polan.Rmd"
output: html_document
---

Today in class, we'll hold a competition for who can best predict an outcome from the consumer expenditure survey. The variable we're interested in today is `grocery`: the amount spent by the family unit on groceries in that quarter. You can use any variables (except `grocery` itself) to predict this outcome. 

For today, we're going to code in pairs. This means that only one person types anything in. The other person sits back and thinks about what should be done and makes suggestions. We'll switch coding pairs halfway through, so everyone will have a chance. 


```{r,echo=FALSE}

library(tidyverse)
library(forcats)
library(ModelMetrics)

load("cex.RData")

## Create training and testing datasets

##---------------------------------------

## This code defines a testing and training dataset
## You don't need it right now, but I want to show you
## how it's done

# Get half of the data via random sample
cex_train<-sample_frac(cex,.5)

save(cex_train,file="cex_train.Rdata")

## Testing data is the other half of the data--- the half of cex not in the 
##training dataset

cex_test<-setdiff(cex,cex_train)

save(cex_test,file="cex_test.Rdata")
##---------------------------------------


##redefine cex as cex training dataset
load("cex_train.Rdata")
cex<-cex_train

my.vals<-names(cex)
explanation<-c("id",
               "education: 10 8th grade, 11 9-12, 12 hs grad,13 some college, 14 assoc, 15 bach 16 masters, 17 prof",
"education of second person",
"urban: 1=urban, 2=not",
"race: 1 white, 2 black, 3 native american, 4 asian 5 pi 6 multi",
"race of other person",
"income class: 1: 0-10k, 2 5-10k, 3 10-15k 4 15 to 20k 5 20-30k 6 30-40k 7 40-50k 8 50-70k 9 70k+",
"inc_rank: percentile rank of income",
"sex of reference: 1 male 2 female",
"sex of other person",
"reference person hispanic? 1 hispanic 2 non",
"other person hispanice? 1 hispanic 2 non",
"poverty: 1 below poverty, 2 not",
"region 1 Ne 2 MW 3 South 4 West",
"family size: ",
"family type 1 hw, 2 hw own children, oldest under 6, 3 hw own children oldest 6-17, 4 hw own children oldest over 17, 5 all other hws, 6 one parent male, kids under 18, 7 one parent female, kids under 18, 8 single, 9 other ",
"Children's ages: 0 no kids, 1 all less 6 2 oldest bt 6 and 11, 3 All children between 6 and 11
 4 Oldest child between 12 and 17 and at least one child less than 12
 5 All children between 12 and 17
 6 Oldest child greater than 17 and at least one child less than 17
 7 All children greater than 17 one under 6",
"quarter and year",
"Dining out",
"Grocery",
"Grocery non-food",
"Grocery food",
"Alcohol at home",
"Alcohol out",
"Other store expenditures",
"Cigarrettes",
"Transportation to work"
)


codebook<-data.frame(my.vals, explanation)



## A bit of wrangling

cex<-cex%>%mutate(pov_cym=ifelse(pov_cym=="",NA,pov_cym))
cex<-cex%>%mutate(pov_cym=fct_recode(as.factor(pov_cym),
                                     "In Poverty"="2",
                                     "Not in Poverty"="3"))
cex<-filter(cex,is.na(pov_cym)==FALSE)

```


```{r model to predict grocery spending}
mod4<-lm(log(grocery+1)~ #log of grocery spending, plus one for zeros
           +pov_cym+ #poverty
           fam_size+ #family size
           bls_urbn+ #whether or not fam is urban
           childage+
           log(other_store+1)+
           log(booze_home+1)+ #drinks booze at home
           fam_type
           #log(dine_out+1)
         ,data=cex_train,na.action="na.exclude")

summary(mod4)

exp(.86)

cex_test<-cex_test%>%mutate(pred4=predict(mod4,newdata=cex_test))

#View(pred4)

compare_data<-data.frame(cex_test$grocery,cex_test$pred4)%>%filter(is.na(cex_test.pred4)==FALSE)

#when you compare data, you have to name the variables based on the dataset they came from (like how pred4 is now cex_test.pred4)

View(compare_data)

rmse_4<-with(compare_data,exp(rmse(log(cex_test.grocery+1),cex_test.pred4)));rmse_4

#this above compares the rmse of the testing and training dataset (becuase the mod4 is based on the train data but the compare data is based on the test data)
```

Here's the process for our competition:

1. When you have a model, run it and calculate the rmse. Let us know when you have a new rmse. 
2. We'll graph the rmse by group in real time. 

3. Be prepared to summarize your model in two ways: first, what variables are included? Second, what types of families are predicted to spend both high amounts and low amounts on groceries? 

3. The winners will get the greatest prize of all: the respect of their peers. 
