Classification: Kaggle Competititon, Part Deux
---

Today, we'll be again working in pairs on a Kaggle competition to predict who gets pizza. Using the pizza dataset, find the best fit to the data. However, to qualify as a winner, you need to do the following: 

```{r libraries}
library(tidyverse)
library(knitr)
library(caret)
library(forcats)
library(tidytext)
library(stringr)
library(AUC)
```
```{r}
za_train<-read.csv("za_train.csv")
save(za_train,file="za_train.Rdata")

za_test<-read.csv("za_test.csv")
save(za_test,file="za_test.Rdata")
```
1. Fit a model using logistic regression using the training dataset (pizza_train.csv).
```{r}
View(za_train)
#Logisitic model

za_test<-za_test%>%mutate(favor=ifelse(grepl(x=request_text,pattern="favor"),1,0))%>% 
  mutate(favor=fct_recode(as.factor(favor),
                "Favor in post"="1",
                "Favor not in post"="0"))

za_train<-za_train%>%mutate(favor=ifelse(grepl(x=request_text,pattern="favor"),1,0))%>% 
  mutate(favor=fct_recode(as.factor(favor),
                "Favor in post"="1",
                "Favor not in post"="0"))

#can see how common the word you're testing is
table(za$favor)
prop.table(table(za$favor))

logit_mod<-glm(got_pizza~ #glm is generalized linear model; the formula works the same way, but we have to specify what family at the end. our outcome is binary so it's binomial, and logit is the type of model
             age+
             karma+
             log(total_posts+1)+
             raop_posts+
             karma+
             student+
             grateful+
             pop_request+ 
             favor+
             score,
             data=za_train,
            na.action=na.exclude,
            family=binomial(link="logit"),
               y=TRUE)

summary(logit_mod)
```


2. Compute the predictions from your model from the testing dataset (pizza_test.csv). 
3. Calculate the AUC (from the Receiver Operating Characteristic) for the predictions from your model from the testing dataset (za_test.csv). 
4. Find a way to plot the predictions from your model.
```{r}
logit_predict<-predict(logit_mod,newdata=za_test,type="response")

logit_predict_no_na<-na.omit(logit_predict)

logit_roc<-roc(logit_predict_no_na,as.factor(logit_mod$y))

auc(logit_roc) #this command (area under curve) tells us the area
 
plot(logit_roc)
```

 

For some ideas, see: http://cs.stanford.edu/~althoff/raop-dataset/altruistic_requests_icwsm.pdf.