---
title: "05-assignment_cook.Rmd"
output:
  word_document: default
  html_document: default
---
 
```{r,echo=FALSE}

library(tidyverse)
library(forcats)
library(ModelMetrics)

load("../lessons/cex.RData")
##need to tell R where to find the data (it's in the lessons file instead of the assignments file)

## Create training and testing datasets

##---------------------------------------

## This code defines a testing and training dataset
## You don't need it right now, but I want to show you
## how it's done

# Get half of the data via random sample
cex_train<-sample_frac(cex,.5)

save(cex_train,file="cex_train.Rdata")

## Testing data is the other half of the data--- the half of cex not in the training dataset

cex_test<-setdiff(cex,cex_train)

save(cex_test,file="cex_test.Rdata")
##---------------------------------------


##redefine cex as cex training dataset
load("cex_train.Rdata")
cex<-cex_train

my.vals<-names(cex)
explanation<-c("id",
               "education: 10 8th grade, 11 9-12, 12 hs grad,13 some college, 14 assoc, 15 bach 16 masters, 17 prof",
"education of second person",
"urban: 1=urban, 2=not",
"race: 1 white, 2 black, 3 native american, 4 asian 5 pi 6 multi",
"race of other person",
"income class: 1: 0-10k, 2 5-10k, 3 10-15k 4 15 to 20k 5 20-30k 6 30-40k 7 40-50k 8 50-70k 9 70k+",
"inc_rank: percentile rank of income",
"sex of reference: 1 male 2 female",
"sex of other person",
"reference person hispanic? 1 hispanic 2 non",
"other person hispanice? 1 hispanic 2 non",
"poverty: 1 below poverty, 2 not",
"region 1 Ne 2 MW 3 South 4 West",
"family size: ",
"family type 1 hw, 2 hw own children, oldest under 6, 3 hw own children oldest 6-17, 4 hw own children oldest over 17, 5 all other hws, 6 one parent male, kids under 18, 7 one parent female, kids under 18, 8 single, 9 other ",
"Children's ages: 0 no kids, 1 all less 6 2 oldest bt 6 and 11, 3 All children between 6 and 11
 4 Oldest child between 12 and 17 and at least one child less than 12
 5 All children between 12 and 17
 6 Oldest child greater than 17 and at least one child less than 17
 7 All children greater than 17 one under 6",
"quarter and year",
"Dining out",
"Grocery",
"Grocery non-food",
"Grocery food",
"Alcohol at home",
"Alcohol out",
"Other store expenditures",
"Cigarrettes",
"Transportation to work"
)


codebook<-data.frame(my.vals, explanation)



## A bit of wrangling

cex<-cex%>%mutate(pov_cym=ifelse(pov_cym=="",NA,pov_cym))
cex<-cex%>%mutate(pov_cym=fct_recode(as.factor(pov_cym),
                                     "In Poverty"="2",
                                     "Not in Poverty"="3"))
cex<-filter(cex,is.na(pov_cym)==FALSE)

View(cex)
```

```{r cig binary variable}
## Crate a binary variable
cex<-cex%>%mutate(cig=ifelse(is.na(cigarettes),0,1))

table(cex$cig)
```  

```{r model to predict cig}
mod<-lm(cig~
           fam_size+
           pov_cym+
          #log(booze_out+1)+
          #log(booze_home+1)+
          ref_race+
          bls_urbn+
           log(grocery+1),
           data=cex,na.action="na.exclude")

summary(mod)

exp(.3681)

cex<-cex%>%mutate(pred=predict(mod))

compare_data<-data.frame(cex$cig,cex$pred)%>%filter(is.na(cex.pred)==FALSE)
names(compare_data)<-c("cig","pred")

rmse<-with(compare_data,exp(rmse(cig,pred)));rmse

```

This model uses the training data (half of the total cex data) to predict cigarette usage with the variables of family size, poverty level, race, grocery spending, and whether or not the family is urban. I originally also has expenditure on alcohol, both in the home and out, as variables but found that they increased the residual error of the model. The RMSE for this model is 1.444 (which can be assessed by taking the exponent of the residual error in the summary or by using the RMSE function). The multiple r-squared shows that this model only accounts for .9% of total prediction for who spends money on cigarettes, so it is still quite ineffective. The t-value is also less than 2, meaning this may actually be equal to zero (there may not be statistical significance). Still, it is the best model I could find based on minimizing RMSE. Part of its inaccuracy is likely because the variable we are trying to predict, cig, is a binary variaable. Regardless, based on the model those in urban famillies and those in poverty have a postive correlation to buying cigarettes (because the correlation coefficient is negative for "not in poverty"). Race2, race4, and race6 have negative correlations to buying cigarettes, and race3, and race5 have postive correlation to buying cigarettes.

```{r new dataset for cigarette buyers}
cex_cig_train<-cex[!(is.na(cex$cigarettes)),]

cex_cig_test<-cex_test[!(is.na(cex$cigarettes)),]
                  
View(cex_cig_train)

```

```{r predict using cex_cig data}
mod1<-lm(log(cigarettes)~
           fam_size+
           pov_cym+
          #log(booze_out+1)+
          #log(booze_home+1)+
          ref_race+
          bls_urbn+
           log(grocery+1),
           data=cex_cig_train,na.action="na.exclude")
#don't need a plus one on the cigarettes log because it's already filtered to people who spent something on cigarettes

summary(mod1)

exp(.7694) 

cex_cig_train<-cex_cig_train%>%mutate(pred1=predict(mod1))

compare_data<-data.frame(cex_cig_train$cigarettes,cex_cig_train$pred1)%>%filter(is.na(cex_cig_train.pred1)==FALSE)
names(compare_data)<-c("cigarettes","pred1")

rmse1<-with(compare_data,exp(rmse(log(cigarettes+1),pred1)));rmse1
 
```
This model is more accurate in predicting who buys cigarettes, likely because it takes into account the amount spent instead of using a binary variable for spending. The RMSE is 2.14, and the multiple r-squared is .07, meaning this still only accounts for 7% of fluctuation within the model. The t value is still under 2, meaning the correlation could equal zero. Negative correlations between the variables and cigarette spending include family size and every race in the model. Poverty, urban families, and grocery spending all have positive correlations. 

```{r rmse with testing data}
#mod1_predict<-predict(mod1,newdata=cex_test)

#exp(rmse(mod1_predict,log(cex_test$cigarettes+1)))

cigs_predict<-predict(mod1,newdata=cex_cig_test)
exp(rmse(cigs_predict,log(cex_cig_test$cigarettes)))
```
