
For this assignment, I want you to use the Census ACS API to download and plot data. Complete the following steps:
```{r}
library(rvest)
library(ggthemes)
library(lubridate)
library(acs)
library(noncensus)
library(leaflet)
library(tigris)
options(tigris_use_cache = TRUE) # save tigris shapefiles
library(plotly)
library(ggplot2)
library(forcats)
library(tidyverse)
```


1. Download data for all of the zip codes in Los Angeles county on education levels. 
2. Compute the proportion of the population that has a bachelor's degree or above by zip code.
3. Download data for all of the zip codes in LA county on family income by zip code. 
4. Compute the proportion of the population that has family income above 75,000. 

```{r}
## Look up fips code for county
lookup_code("CA","Los Angeles") 

state_fips<-"06"
county_stub<-"037"
county_fips<-paste0(state_fips,county_stub)

# Get dataset that matches all zip codes to cities, counties and states. 
county_to_zip<-read_csv("http://www2.census.gov/geo/docs/maps-data/data/rel/zcta_county_rel_10.txt")

#easier names to work with
names(county_to_zip)<-tolower(names(county_to_zip))

#Just zip codes in selected county
county_to_zip<-county_to_zip%>%filter(state==state_fips,county==county_stub)%>%
  select(zcta5,state,county)

#list of zip codes
ziplist<-county_to_zip$zcta5

#City names
data(zip_codes)

city_zip<-zip_codes%>%filter(zip%in%ziplist)%>%select(zip,city)

#Arrange in order
city_zip<-city_zip%>%arrange(as.numeric(zip))

city_zip
```
```{r}
#my acs key
acs_key<-"6a40376acbf5ba1669c71e128f22503adba853c8"

#List of tables: https://www.census.gov/programs-surveys/acs/technical-documentation/summary-file-documentation.html under, 1-year appendices
# b15002: education of pop over 25, by sex 
# b19001: household income over last 12 months

get_acs_data<-TRUE #set this up so that after you get the data you can turn this to "false" and not have to run this every time cuz it takes awhile

if (get_acs_data==TRUE){
  
api.key.install(acs_key, file = "key.rda")

select_zip<-geo.make(zip.code=ziplist) 

county_educ=acs.fetch(geography=select_zip, 
                      endyear=2014,
                      table.number="B15002",
                      col.names="pretty",verbose=T) 

acs.colnames(county_educ)

## Proportion of individuals with or above Bachelor's= number with college degree/ total number
prop_coll_above<-divide.acs(numerator=(county_educ[,15]+
                                      county_educ[,16]+
                                      county_educ[,17]+
                                      county_educ[,18]+ 
                                      county_educ[,32]+
                                      county_educ[,33]+
                                      county_educ[,34]+
                                      county_educ[,35]),
                            denominator=county_educ[,1]
)

                            
# 19001-- family income           
county_income<-acs.fetch(geography=select_zip, 
                        endyear = 2015,
                        table.number="B19001", 
                        col.names="pretty")

acs.colnames(county_income)

#Proportion above 75k-- 
prop_above_75<-divide.acs(numerator=(county_income[,13]+
                            county_income[,14]+
                            county_income[,15]+
                            county_income[,16]+
                            county_income[,17]),
                          denominator=county_income[,1]
                          )
                          
# Convert to tibble
county_df<-tibble(substr(geography(county_educ)[[1]],7,11),
                       as.numeric(estimate(prop_coll_above)),
                       as.numeric(estimate(prop_above_75))
) #keep in mind that these are estimates and not exact, so it's worth checking if you're working with a really small geographical area (not as big a worry if it's a big area)

# Give it easy to use names
names(county_df)<-c("zip","college_educ","income_75")
save(county_df,file="dav.RData")
}else{
  load("dav.RData")
}

View(county_df) 
```


5. Download data for all of the zip codes in LA county on health insurance coverage status (you'll need to look that table up here:[http://www.census.gov/programs-surveys/acs/technical-documentation/summary-file-documentation.html](http://www.census.gov/programs-surveys/acs/technical-documentation/summary-file-documentation.html).

6. Calculate the proportion of the population in each zip code that is uninsured. 
```{r}
#my acs key
acs_key<-"6a40376acbf5ba1669c71e128f22503adba853c8"

get_acs_data<-TRUE #set this up so that after you get the data you can turn this to "false" and not have to run this every time cuz it takes awhile

if (get_acs_data==TRUE){
  
api.key.install(acs_key, file = "key.rda")

select_zip<-geo.make(zip.code=ziplist) 

county_insur=acs.fetch(geography=select_zip, 
                      endyear=2014,
                      table.number="B27001", #health insurance by sex by age
                      col.names="pretty",verbose=T) 

acs.colnames(county_insur)

## Proportion of individuals without insurance (uninsured all ages male and female over total)
prop_uninsured<-divide.acs(numerator=(county_insur[,5]+
                                      county_insur[,8]+
                                      county_insur[,11]+
                                      county_insur[,14]+ 
                                      county_insur[,17]+
                                      county_insur[,20]+
                                      county_insur[,23]+
                                      county_insur[,26]+
                                      county_insur[,29]+
                                      county_insur[,33]+
                                      county_insur[,36]+
                                      county_insur[,39]+
                                      county_insur[,42]+
                                      county_insur[,45]+
                                      county_insur[,48]+
                                      county_insur[,51]+
                                      county_insur[,54]+
                                      county_insur[,57]),
                            denominator=county_insur[,1]
)


# Convert to tibble
county_un<-tibble(substr(geography(county_insur)[[1]],7,11),
                       as.numeric(estimate(prop_uninsured))
) 

# Give it easy to use names
names(county_un)<-c("zip","uninsured")
save(county_un,file="insur.RData")
}else{
  load("insur.RData")
}

View(county_un) 
```


7. Plot the proportion uninsured as a function of education, and then as a function of income. 
```{r education}
#Merge  datasets
zip_df_un<-left_join(county_df,county_un,by="zip")
zip_df_un<-left_join(zip_un,city_zip,by="zip")

#Bar plot of uninsured by education
g6<-ggplot(zip_df_un,
           aes(x=fct_reorder(as.factor(college_educ),uninsured),y=uninsured,fill=uninsured))

g6<-g6+scale_fill_gradient(low="blue",high="red")

g6<-g6+geom_bar(stat="identity")

g6<-g6+xlab("Proportion of Residents with College Degree")+ylab("Proportion of Uninsured")

g6<-g6+theme(legend.position="none")

g6

```
```{r income}
#Bar plot of uninsured by income
g6<-ggplot(zip_df_un,
           aes(x=income_75,y=uninsured,fill=uninsured))

g6<-g6+scale_fill_gradient(low="blue",high="red")

g6<-g6+geom_bar(stat="identity")

g6<-g6+xlab("Proportion of Residents with Income>75K")+ylab("Proportion of Uninsured")

g6<-g6+theme(legend.position="none")

g6

```
8. Model the proportion uninsured as a function of education, income _and one other variable of your choice_. 

```{r}
#get data about employment status in these zip codes
acs_key<-"6a40376acbf5ba1669c71e128f22503adba853c8"

get_acs_data<-TRUE #set this up so that after you get the data you can turn this to "false" and not have to run this every time cuz it takes awhile

if (get_acs_data==TRUE){
  
api.key.install(acs_key, file = "key.rda")

select_zip<-geo.make(zip.code=ziplist) 

county_empl=acs.fetch(geography=select_zip, 
                      endyear=2014,
                      table.number="B27011", #health insurance by employment status
                      col.names="pretty",verbose=T) 

acs.colnames(county_empl)

# Convert to tibble
county_job<-tibble(substr(geography(county_empl)[[1]],7,11)
) #the 7-11 is where the zipcode is (so that R reads the zipcode and not all the other stuff)

# Give it easy to use names
names(county_job)<-c("zip","employment")
save(county_job,file="job.RData")
}else{
  load("job.RData")
}

View(county_job) 

zip_df_un<-left_join(zip_df_un,county_job,by="zip")

#Predict uninsured by income and education and employment status
mod1<-lm(uninsured~college_educ+income_75+county_empl,data=zip_df_un);summary(mod1)

```


## COMMENTS from WD

## Basic idea is right, but you needed to calculate proportions in the
## last step, doesn't look like that's what happened.