Assignment 8
---

"When life gives you lemons, don’t make lemonade. Make life take the lemons back! Get mad!" -Cave Johnson

For this assignment, you'll be using the lemons dataset, which is a subset of the dataset used for a Kaggle competition described here: 
https://www.kaggle.com/c/DontGetKicked/data. Your job is to predict which cars are most likely to be lemons. 

Complete the following steps.
**you don't have to do all the wrangling or the sentiment analysis. Just table dependent variable and table depedent variable by make, run a linear model, run an AUC on the linear model, run a logistic regression and calculate AUC for that logit model. No testing and training datasets either. 

```{r libraries}
library(tidyverse)
library(knitr)
library(caret)
library(forcats)
library(tidytext)
library(stringr)
library(AUC)
```

```{r data}
lem_test<-read_csv("lem_test.csv")
lem_train<-read_csv("lem_train.csv")

View(lem_train)
```
1. Calculate the proportion of lemons in the training dataset using the `IsBadBuy` variable. (table)
```{r}
lem_train%>%
  count(IsBadBuy)%>% 
  mutate(p=prop.table(n))%>% #mutate for proportions using prop.table
  kable(format="markdown") # output to table

#12.3% of vehicles in the training dataset are lemons
```
2. Calculate the proportion of lemons by Make. (proportions table by make)
```{r by make}
#Predictions using conditional means

lem_train%>%group_by(Make)%>%summarize(mean(IsBadBuy))
```


3. Now, predict the probability of being a lemon using a linear model (`lm(y~x`), with covariates of your choosing from the training dataset. 
```{r linear_model}
# Linear model

lm_mod<-lm(IsBadBuy~
             VehYear+
             VehicleAge+ 
             Make+ 
             Nationality+
             WarrantyCost,
           data=lem_train,y=TRUE,na.exclude=TRUE);summary(lm_mod)
```

4. Make predictions from the linear model.
```{r}
#Predictions

## Linear model predictions
lm_predict<-predict(lm_mod)

## Convert to binary, 1= >.5
lm_predict_bin<-ifelse(lm_predict>.1,1,0) #made it .1 because my model was incredibly sensitive but not very specific at first (was classifying everyone as a 1 when the prediction is above a .1)

lm_predict_bin 

## Table of actual vs. predicted, what's going on here?
lm_table<-table(lm_predict_bin,lm_mod$y)

## Percent correctly predicted
pcp<-(lm_table[1,1]+lm_table[2,2])/sum(lm_table)

pred_table<-prop.table(lm_table,margin=1)

pred_table

rownames(pred_table)<-c("Predicted: Yes","Predicted: No")
colnames(pred_table)<-c("Actual: Yes","Actual: No")

## Generate confusion matrix
confusionMatrix(data=lm_predict_bin,
                reference = lm_mod$y,positive="1")
```

5. Calculate the AUC for the linear predictions from the ROC against the outcome for the training dataset. 
```{r}
lm_predict_no_na<-na.omit(lm_predict)

lm_roc<-roc(lm_predict_no_na,as.factor(lm_mod$y))

auc(lm_roc) #this command (area under curve) tells us the area
 
#the auc of my linear model is .65
plot(lm_roc)
```

6. Now, predict the probability of being a lemon using a logistic regression (`glm(y~x,family=binomial(link="logit")`)), again using covariates of your choosing. Add these to the existing linear model already give to you.  
```{r}
#Logisitic model

logit_mod<-glm(IsBadBuy~ 
                VehYear+
             VehicleAge+ 
             Make+ 
             Nationality+
             WarrantyCost,
             data=lem_train,
            na.action=na.exclude,
            family=binomial(link="logit"),
               y=TRUE)

summary(logit_mod)
```
```{r linear_model2}
# Second Linear model

lm_mod<-lm(IsBadBuy~
             VehYear+
             VehicleAge+ 
             Make+ 
             Nationality+
             logit_predict+ #added the logit_predict variable in
             WarrantyCost,
           data=lem_train,y=TRUE,na.exclude=TRUE);summary(lm_mod)
```

7. Make predictions from the logit model. Make sure these are probabilities. 
```{r}
logit_predict<-predict(logit_mod,type="response")
logit_predict
```

8. Calculate the AUC for the linear predictions from the ROC. 
```{r}
logit_predict_no_na<-na.omit(logit_predict)

logit_roc<-roc(logit_predict_no_na,as.factor(logit_mod$y))

auc(logit_roc) 
#the auc for my logit model is .65 also

plot(logit_roc)
```
9. (optional) submit your predictions from the testing dataset as a late submission to Kaggle and see how you do against real-wolrd competition. 

## COMMENTS from WD

## Good, nice work!