Assignment 9
---

Complete the following steps:

1. Using the nyclfights13 library, open all of the data tables and turn them into a databse. Make sure to include flights, airlines, airports, weather and planes.
```{r}
# Mostly taken from : http://cran.r-project.org/web/packages/dplyr/vignettes/databases.html circa 2014

#Get libraries
library(tidyverse)
library(DBI)
library(dbplyr)
library(RSQLite)
library(nycflights13)
library(Lahman)

# Data sources from NYC flights
data(flights)
data(airlines)
data(airports)
data(weather)
data(planes)

con <- dbConnect(RSQLite::SQLite(), ":memory:")

#Write flights tables to database (you won't usually do this)
dbWriteTable(con,
             "flights", 
             as.data.frame(flights))

dbWriteTable(con,
             "planes", 
             as.data.frame(planes))

dbWriteTable(con,
             "airlines", 
             as.data.frame(airlines))

dbWriteTable(con,
             "weather", 
             as.data.frame(weather))

dbWriteTable(con,
             "airports", 
             as.data.frame(airports))

#List tables in database
dbListTables(con)

#List fields in various tables
dbListFields(con,"airlines")

dbListFields(con,"planes")

dbListFields(con,"flights")

#struggling to find a way to view the column names in a particle table in the database (trying to see what the ent)
Columns <- as.data.frame(colnames(dbFetch(con, "flights")))
```

2. Create a data frame from a SQL pull from the database that consists only of flights that took off from JFK in May.
```{r}
#New request: which carriers have longer delays? 
req1_text<-"SELECT a.carrier, a.name, f.origin, f.month
            FROM flights f 
            JOIN airlines a ON a.carrier=f.carrier 
            WHERE f.month=5 AND f.origin=JFK
            "

#doesn't work because says that JFK column doesn't exist in the f.origin, so I thought it could be numeric like the months are, but that came up with 0 results (which is highly unlikely that there were no flights from JFK in May)
req1<-dbSendQuery(con,req1_text)

req1_df<-dbFetch(req1,n=-1)

dbClearResult(req1)

head(req1_df) 

```

3. Create a data frame from a SQL pull from the database that consists only of flights that took off on-time (a delay of less than 10 minutes) from Newark at temperatures of less than 40 degrees F.

```{r}
#SQL request, flight info combined with weather info
weather_text<-"SELECT f.dep_delay, f.origin, w.temp
              FROM weather w
              JOIN flights f
              ON f.year=w.year
              AND f.month=w.month
              AND f.day=w.day
              WHERE f.dep_delay<10 AND w.temp<40 AND f.origin=NWK"

weather_req<-dbSendQuery(con,weather_text)

weather_df<-dbFetch(weather_req,n=-1)

dbClearResult(weather_req)

head(weather_df)
```

4. Create data frame from a SQL pull from the database that consists of planes flown by United.
```{r}
req_text<-"SELECT a.carrier, a.name, f.tailnum
            FROM flights f
            JOIN airlines a 
            ON a.carrier=f.carrier 
            WHERE a.carrier=UA
            "

req<-dbSendQuery(con,req_text)

req_df<-dbFetch(req,n=-1)

dbClearResult(req)

head(req_df) 

```

5. Create a model that tests whether certain plane types (e.g. 737s) are more likely to have arrival delays than other planes. You can include other variables in the model. You'll need to do a pull from more than one table to get the data you need. 
```{r}
req2text<-"SELECT f.dep_delay, p.tailnum, f.carrier, p.type, p.model FROM planes p JOIN flights f ON p.tailnum=f.tailnum"

req2<-dbSendQuery(con,req2text)

req2_df<-dbFetch(req2,n=-1)

dbClearResult(req2)

head(req2_df) 

mod<-lm(dep_delay~
           tailnum+
           carrier+
           model+
           type,
           data=req2_df)

summary(mod)


```



## COMMENTS from WD

## Good, nice work!