---
title: "assignment3_cook.Rmd"
output:
  word_document: default
  html_document: default
---

Load packages and attrition dataset
```{r  include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(forcats)
library(RColorBrewer)
```
```{r}

load("../lessons/at.Rdata")
View(at)
```
```{r}
#so that I hide everything except graphics and actual text in the knit version
echo=FALSE 
results='hide' 
warning=FALSE 
message=FALSE
```
1.  Distribution of monthly income.
```{r}
gg<-ggplot(at,aes(x=MonthlyIncome))
gg<-gg+geom_histogram(binwidth = 200,fill="lightblue")
gg
## Axes Labelling 
gg<-gg+xlab("Monthly Income")+ylab("Number of Employees")
gg
``` 

2.  Average level of monthly income by field of education.
```{r}

at_sum<-at%>%
  group_by(EducationField)%>%
  summarize(MonthlyIncome=mean(MonthlyIncome))

at_sum

```
```{r}
## Bar plot with monthly income as height and education field as category, bars ordered by level
gg<-ggplot(at_sum,aes(x=fct_reorder(EducationField,MonthlyIncome),y=MonthlyIncome))

gg<-gg+geom_bar(stat="identity")

## Labeling
gg<-gg+xlab("Education Field")+ylab("Monthly Income")
##Print
gg
```

3.  Average level of monthly income by field of education and gender.
```{r}
## Summarize income by field of education and gender
at_sum<-at%>%
  group_by(EducationField,Gender)%>%
  summarize(MonthlyIncome=mean(MonthlyIncome))

## Get the results
at_sum

## Plot it using a bar plot
gg<-ggplot(at_sum,aes(x=reorder(EducationField,MonthlyIncome),y=MonthlyIncome,color=Gender))
gg<-gg+geom_bar(stat="identity",aes(fill=Gender),position="dodge")
gg<-gg+xlab("Education Field")+ylab("Monthly Income")
gg
```

4. Average levels of monthly income by field of education, gender and job level (scale of 1-5, highest ranked employees are 5)
```{r}

at_sum<-at%>%
  group_by(EducationField,Gender,JobLevel)%>%
  summarize(MonthlyIncome=mean(MonthlyIncome))

at_sum

at_sum%>%arrange(MonthlyIncome)

gg<-ggplot(at_sum,aes(x=reorder(EducationField,MonthlyIncome),
                      y=MonthlyIncome,
                      fill=Gender))
## Bar plot, with unstacked (dodge)
 gg<-gg+geom_bar(stat="identity",position="dodge")
## Separate out by job level
gg<-gg+facet_wrap(~JobLevel)
## Change orientation to sideways
gg<-gg+coord_flip()
gg<-gg+xlab("Education Field")+ylab("Monthly Income, Grouped by Job Level 1-5")
## Print
gg
```

5. Average levels of montly income by field of education, gender, job level and department.
```{r ,fig.height=6}

at_sum<-at%>%
  group_by(EducationField,Gender,JobLevel,Department)%>%
  summarize(MonthlyIncome=mean(MonthlyIncome))%>%
  ungroup()%>%
  arrange(MonthlyIncome)

at_sum

## A new variable for every type of employee defined by travel, gender, marital status, and happiness with work/life

at_sum$grouping<-paste0(at_sum$EducationField,
                       ", ",
                       at_sum$Gender,
                       ", ",
                       at_sum$JobLevel,
                       ", Work/Life:",
                       at_sum$Department)

at_sum$grouping<-as.factor(at_sum$grouping)

at_sum%>%select(grouping,MonthlyIncome)

at_sum<-at_sum%>%filter(MonthlyIncome>.01)

gg<-ggplot(at_sum,aes(x=fct_reorder(grouping,MonthlyIncome),y=MonthlyIncome))
gg<-gg+geom_bar(stat="identity",aes(fill=Department))
gg<-gg+coord_flip()
gg<-gg+xlab("Education Field, Gender, Job Level")+ylab("Monthly Income")
gg

```


