---
title: "02-progress_report_cook.Rmd"
output: html_document
---
 
```{r  include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)
library(gridExtra)
library(ModelMetrics)
library(haven)
library(readxl)
library(forcats)
library(RColorBrewer)
```

```{r coal state emissions}
download.file("https://www.eia.gov/environment/emissions/state/excel/coal_CO2_by_state_2014.xlsx",destfile = "coal_co2.xlsx",method="curl")

coal_co2<-read_excel("coal_co2.xlsx")

coal_co2

View(coal_co2)
```
Dependent Variable: CO2 Emissions by State
Graphic: Rank of State Percent Change in CO2 Emissions from 1980-2014
```{r univariate plot}
## Create a rank variable for percent change (excluding unnecessary rows)

coal_co2_slim<-coal_co2[-(c(1,2,11,54:58)), ]

View(coal_co2_slim)

coal_co2_slim<-coal_co2_slim%>%mutate(percent_change_rank=rank(X__36))

##should figure out how to make a bar chart with states on x axis, ordered by rank of percent change, and the percent change stat on the y axis

#want to do a reorder so that X__36 is ranked with 1 being the lowest change (decreased emissions) and 50 being a high, positive number of percent change emissions
#want to plot that reorder but use the actual state names instead of the order of the states based on the variable (idk how to change the names)
#want to change the names of the column titles in my excel data (so I can left-join later on)
#want to have the y axis data (percent change) done in increments so it's all legible, instead of having every number displayed so that it's all crammed in there

coal_co2_slim$X__36 <- factor(coal_co2_slim$X__36,levels=coal_co2_slim$X__36[order(coal_co2_slim$X__36)])

fct_reorder(coal_co2_slim$X__36)

coal_co2_slim$X__36  

## Plot by rank
gg<-ggplot(data=coal_co2_slim , aes(x=column1, #need to figure out how to say column 1 aka the state names; also need to rename the states to their abbreviation
                         y=X__36)) +geom_bar(stat="identity")
##Add Axis Labels
gg<-gg+xlab("State Rank")+ylab("Percent Change 1980-2014")

## Add Points
gg<-gg+geom_point(alpha=1,size=.5)

gg
```

Conditional Mean of 2014 State CO2 Emissions as a Function of Emission Cap Laws  
```{r Conditional Mean}
##Condtional Average across a single variable

#open csv file I created with binary for co2 emission cap laws by state
cap_laws<-read_csv(file="emission_caps.csv")

head(cap_laws)

View(cap_laws)

#trying to merge my binary csv with coal data but won't let me because the labels are different
#need to figure out how to change labels
coal_laws_total<-merge(cap_laws,coal_co2,by=State)
#trying to group states by binary (whether or not they have the law)
#is there a way to group by response (as in 0 or 1) instead of rank?
cap_laws_group<-cap_laws%>%mutate(binary_rank=rank(X2))

#trying to use this binary grouping as a predictor for co2 emissions, but struggling because can't figure out how to cross between 2 datasets in the commands (X__35 is the 2014 emissions data from the coal dataset but the binary data is in a separate csv file)
cap_laws_group<-cap_laws_group%>%group_by(binary_rank)%>%
coal_co2_slim%>%mutate(pred_emissions_binary=mean(X__35))

  ungroup()%>% 
  #Rank by prediction, with ties sorted randomly
  mutate(pred_emissions_binary_rank=rank(pred_emissions_binary,ties.method="random"))

  
#graphing 2014 emissions as a function of emission cap laws  
gg<-ggplot(data=coal_co2_slim,aes(x=pred_emissions_binary_rank,y=X__35,color="Actual"))
gg<-gg+geom_point(alpha=.5,size=.5)
gg<-gg+geom_point(aes(x=pred_emissions_binary_rank,y=pred_emissions_binary,color="Predicted:Conditional Mean, 1 var"))
gg<-gg+ scale_color_manual("Type",values=c("Predicted:Conditional Mean, 1 var"="red","Actual"="black"))
gg<-gg+theme(legend.position="bottom")
gg<-gg+xlab("Rank")+ylab("2014 State Emissions")
gg  
  
```




 COMMENTS from WD

 Okay, looks like there's some confusion here. Recoding to 0,1 is done
 using mutate(ifelse . . .) . The key graphic you're looking for is a
 barplot of emissions by regulations, yes? This graphic is not what you
 want. You need to use left_join() to put the two datasets together,
 but they need a common id variable, such as state name. The variable
 for state name has to have the same name in each dataset. Check out
 the Wickham book, the chapter on relational data, section 13.4 on
 joins: http://r4ds.had.co.nz/relational-data.html#mutating-joins This
 will show you how to combine two different datasets.