```{r  include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)
library(gridExtra)
library(ModelMetrics)
library(haven)
library(readxl)
library(forcats)
library(RColorBrewer)
```

For this progress report, I need the following in an .Rmd file:

1. Code that accesses a dataset that contains your dependent variable and most of your relevant independent variables. 

```{r coal state emissions}
download.file("https://www.eia.gov/environment/emissions/state/excel/coal_CO2_by_state_2014.xlsx",destfile = "coal_co2.xlsx",method="curl")

coal_co2<-read_excel("coal_co2.xlsx")

View(coal_co2)

## Create a rank variable for percent change (excluding unnecessary rows)

coal_co2_slim<-coal_co2[-(c(1,2,11,54:58)),-(c(2:36))]
coal_co2_slim<-coal_co2_slim%>%mutate(percent_change_rank=rank(`X__36`))

colnames(coal_co2_slim)<-c("State","Percent Change CO2 Emissions 1980-2014","Absolute Change CO2 Emissions 1980-2014","Percent Change Rank")

View(coal_co2_slim)

#open csv file I created with binary for co2 emission cap laws by state
cap_laws<-read_csv(file="emission_caps.csv")

colnames(cap_laws)<-c("State","Emission Cap Laws")
cap_laws<-cap_laws[-(c(1)), ]

head(cap_laws)

View(cap_laws)

coal_laws_total<-merge(cap_laws,coal_co2_slim,by="State")

View(coal_laws_total)
#trying to group states by binary (whether or not they have the law)
#is there a way to group by response (as in 0 or 1)?

#once i figure this out, can do the same thing with multiple laws using a csv file of binary code for each state
```
1. A univaraiate graphic describing the dependent variable.
 
Dependent Variable: CO2 Emissions by State
Graphic: Rank of State Percent Change in CO2 Emissions from 1980-2014
----is this the best graphic I can do? Should I focus in on a particular year of change or a particular year of emissions so that it's more focused on potential policy enactment?
```{r univariate plot}

##should figure out how to make a bar chart with states on x axis, ordered by rank of percent change, and the percent change stat on the y axis (percent change will assumably control for GDP and population)

#want to do a reorder so that percent change is ranked with 1 being the lowest change (decreased emissions) and 50 being a high, positive number of percent change emissions
#want to plot that reorder but use the actual state names instead of the order of the states based on the variable (idk how to change the names)
#want to change the names of the column titles in my excel data (so I can left-join later on)
#want to have the y axis data (percent change) done in increments so it's all legible, instead of having every number displayed so that it's all crammed in there

fct_reorder(coal_laws_total$`Percent Change Rank`,na.rm=TRUE)

## Plot by rank
gg<-ggplot(data=coal_laws_total , aes(x=State, #need to figure out how to say column 1 aka the state names; also need to rename the states to their abbreviation
                         y='Percent Change CO2 Emissions 1980-2014') +geom_bar(stat="identity"))
##Add Axis Labels
gg<-gg+xlab("State by Rank")+ylab("Percent Change 1980-2014")

## Add Points
gg<-gg+geom_point(alpha=1,size=.5)

gg
```

1. Results of two different analyses that predict your dependent variable (conditional means, regression, classification, k-means clustering)

```{r}
#can run regression on change in state emission level as a function of whether or not they have 2 different laws and then plot like i did for assignment 6
#can even create a variable that's based on number of laws each state has total regarding co2 emission (so like if California has 30 laws, that'll be their observation etc etc)

```
1. Two graphics that show the results of your analyses above. 
