```{r  include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)
library(gridExtra)
library(ModelMetrics)
library(haven)
library(readxl)
library(forcats)
library(RColorBrewer)
library(rvest)
```

For this progress report, I need the following in an .Rmd file:

1. Code that accesses a dataset that contains your dependent variable and most of your relevant independent variables. 

```{r coal state emissions}
download.file("https://www.eia.gov/environment/emissions/state/excel/coal_CO2_by_state_2014.xlsx",destfile = "coal_co2.xlsx",method="curl")

coal_co2<-read_excel("coal_co2.xlsx")

View(coal_co2)

## Create a rank variable for percent change (excluding unnecessary rows and all columns that aren't percent change)

coal_co2_slim<-coal_co2[-(c(1,2,11,54:58)),-(c(2:36))]
coal_co2_slim<-coal_co2_slim%>%mutate(percent_change_rank=rank(`X__36`))

colnames(coal_co2_slim)<-c("state","percent_change_emissions","absolute_change_emissions","percent_change_rank")

View(coal_co2_slim)

#open csv file I created that has: abbreviation of state name, binary for co2 emission cap laws by state, and investment in renewable energy by state (in USD)
cap_laws<-read_csv(file="emission_caps.csv")

colnames(cap_laws)<-c("state","abbreviation","emission_cap","renewable_investment")
cap_laws<-cap_laws[-(c(1)), ]

View(cap_laws)

coal_laws_total<-merge(cap_laws,coal_co2_slim,by="state")

coal_laws_total>-coal_laws_total%>%mutate(percent_change_emissions=as.numeric(percent_change_emissions)) #this is to recode the emissions variable so that it's numeric (since the next code chunck has an error message about non-numeric variables)... however, it keeps sending an error message about an invalid argument to unary operator

View(coal_laws_total)
```
1. A univaraiate graphic describing the dependent variable.
 
Dependent Variable: CO2 Emissions by State
Graphic: Rank of State Percent Change in CO2 Emissions from 1980-2014
```{r univariate plot}

#create a new variable that is my percent change rank of state in order (1 being the least increase (aka biggest decrease) in c02 emission and 50 being the biggest increase in emissions)
coal_laws_total>-coal_laws_total%>%mutate(state_sort=fct_reorder(f=coal_laws_total,x=percent_change_rank),ties.method="random")

coal_laws_total>-coal_laws_total%>%mutate(state_sort=fct_reorder(percent_change_rank),ties.method=random)


#this is the code for the plot, but it won't let me actually run the code because it says I have a non-numeric argument to the binary operator; I originally assumed it doesn't think the percent change variable is numeric (since I know for sure the rank variable is numeric), but I tested with other variables and get the same error message. Either way, I'm unsure how to change that since I pulled the emission numbers off the EPA website and can't fix them like I could the variables I made in excel                                          
## Plot by rank
gg<-ggplot(data=coal_laws_total,aes(x=state_sort,y=percent_change_emissions) +geom_bar())
##Add Axis Labels
gg<-gg+xlab("State by Rank of Change in CO2 Emissions")+ylab("CO2 Emissions Percent Change 1980-2014")

gg<-coord_flip()

#add breaks for the emissions data instead of having every numeral from all 50 states
gg<-gg+scale_y_continuous(breaks=c(-7,-6,-5,-4,-3,-2,-1,0,1,2,3,4,5,6,7,8))
## Add Points
gg<-gg+geom_point(alpha=1,size=.5)

gg
```

1. Results of two different analyses that predict your dependent variable (conditional means, regression, classification, k-means clustering)

```{r}

#this is the code for the regression and conditional average, but it won't let me run these either because it says there are non-numeric arguments (same error message as plot)
#regression of percent change of emissions by state based on that state's investment in renewable energy and the existence of legal state-wide co2 emission caps
mod<-lm(percent_change_emissions~ 
           log(renewable_investment)+ 
           emission_cap,
         data=coal_laws_total,na.action="na.exclude")
summary(mod)

#create a variable of the predictions of the regression
coal_laws_total<-coal_laws_total%>%mutate(pred=predict(mod))

##Condtional Average of percent change in co2 emissions based on 4 levels of investment in renewable energy by state
## Create a variable for quartiles of investment in renewable energy
coal_laws_total<-coal_laws_total%>%mutate(investment_level=ntile(renewable_investment,4,na.action="na.exclude"))

table(coal_laws_total$investment_level)

coal_laws_total<-coal_laws_total%>%group_by(investment_level)%>% ## Group by predictor
  ##Calculate mean at each level of predictor
  mutate(pred_emission_investment=mean(percent_change_emissions))%>% 
  ## Ungroup
  ungroup()%>% 
  #Rank by prediction, with ties sorted randomly
  mutate(pred_emission_investment_rank=rank(pred_emission_investment,ties.method="random"))

```
1. Two graphics that show the results of your analyses above. 

```{r}

#scatterplot of percent change of emissions based on investment level, with whether or not emission laws exist as the color
g1<-ggplot(data=coal_laws_total, aes(x=investment_level,y=percent_change_emissions,color=emission_cap)+
           geom_point(shape=1)+ #specify points
           geom_smooth(method=lm)) #ask for lm line
g1


#scatterplot of actual percent change emissions versus predicted percent change emissions from the regression
g2<-ggplot(coal_laws_total, aes(x=percent_change_emissions,y=exp(pred))
+geom_point(shape=1))
g2

#scatterplot of conditional mean
gg<-ggplot(data=coal_laws_total,aes(x=pred_emission_investment_rank,y=percent_change_emissions,color="Actual"))
gg<-gg+geom_point(alpha=.5,size=.5)
gg<-gg+geom_point(aes(x=pred_emission_investment_rank,y=pred_emission_investment,color="Predicted:Conditional Mean, 1 var"))
gg<-gg+theme(legend.position="bottom")
gg<-gg+xlab("Rank")+ylab("Percent Change CO2 Emissions")
gg

```

