---
title: "Project Data"
output: html_document
---

```{r} 
library(tidyverse)
library(haven)
library(readxl)
```

```{r total emissions}
##https://www.epa.gov/sites/production/files/2016-08/us-ghg-emissions_fig-1.csv

if(file.exists("emissions.csv")==FALSE){
  download.file("https://www.epa.gov/sites/production/files/2016-08/us-ghg-emissions_fig-1.csv",destfile="emissions.csv")
emissions<-read_csv("emissions.csv")  
}else{
  emissions<-read_csv("emissions.csv")
}

emissions<-emissions[-(c(1:5)), ]

View(emissions)
```

```{r industry}
if(file.exists("industry.csv")==FALSE){
  download.file("https://www.epa.gov/sites/production/files/2016-08/us-ghg-emissions_fig-2.csv",destfile="industry.csv")
industry<-read_csv("industry.csv")  
}else{
 industry<-read_csv("industry.csv")
}

industry<-industry[-(c(1:5)), ]

View(industry)
```

```{r GDP}
if(file.exists("gdp.csv")==FALSE){
  download.file("https://www.epa.gov/sites/production/files/2016-08/us-ghg-emissions_fig-3.csv",destfile="gdp.csv")
gdp<-read_csv("gdp.csv")  
}else{
 gdp<-read_csv("gdp.csv")
}

gdp<-gdp[-(c(1:5)), ]

View(gdp)

#trying to find explanations for the drop in total US emissions in 2009; it was the year Obama won the presidency, with the support of numerous environmental groups. it was also the year he enacted a program that set the first-ever standard for greenhouse gas emissions for cars
```

timeline of all federal environmental policy: http://www.naem.org/?CP_COMP_milestones

table of different types of policies and how long states have had them: http://www.nature.com/nclimate/journal/v4/n11/fig_tab/nclimate2385_T1.html (could make a binary chart for each state and run a regression)

```{r per capita state emission}
#https://www.eia.gov/environment/emissions/state/analysis/excel/table5.xlsx

if(file.exists("pcstate.xls")==FALSE){
  download.file("https://www.eia.gov/environment/emissions/state/analysis/excel/table5.xlsx",destfile="pcstate.xls")
pcstate<-read_excel("pcstate.xls")  
}else{
 pcstate<-read_excel("pcstate.xls")
}

View(pcstate)
```

take a look at the APIs and how we can use

can have multiple binary variables for the bivariate graphic NOA data (national oceanic something data online) and show different graphics of what's the average contaminants for states with this policy versus this policy etc


