---
title: "Project Data"
output: html_document
---
```{r  include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)
library(gridExtra)
library(ModelMetrics)
library(haven)
library(readxl)
library(forcats)
library(RColorBrewer)
```


```{r total emissions}
##https://www.epa.gov/sites/production/files/2016-08/us-ghg-emissions_fig-1.csv

if(file.exists("emissions.csv")==FALSE){
  download.file("https://www.epa.gov/sites/production/files/2016-08/us-ghg-emissions_fig-1.csv",destfile="emissions.csv")
emissions<-read_csv("emissions.csv")  
}else{
  emissions<-read_csv("emissions.csv")
}

emissions<-emissions[-(c(1:5)), ]

View(emissions)
```

```{r industry}
if(file.exists("industry.csv")==FALSE){
  download.file("https://www.epa.gov/sites/production/files/2016-08/us-ghg-emissions_fig-2.csv",destfile="industry.csv")
industry<-read_csv("industry.csv")  
}else{
 industry<-read_csv("industry.csv")
}

industry<-industry[-(c(1:5)), ]

View(industry)
```

```{r GDP}
if(file.exists("gdp.csv")==FALSE){
  download.file("https://www.epa.gov/sites/production/files/2016-08/us-ghg-emissions_fig-3.csv",destfile="gdp.csv")
gdp<-read_csv("gdp.csv")  
}else{
 gdp<-read_csv("gdp.csv")
}

gdp<-gdp[-(c(1:5)), ]

View(gdp)

#trying to find explanations for the drop in total US emissions in 2009; it was the year Obama won the presidency, with the support of numerous environmental groups. it was also the year he enacted a program that set the first-ever standard for greenhouse gas emissions for cars
```

timeline of all federal environmental policy: http://www.naem.org/?CP_COMP_milestones

table of different types of policies and how long states have had them: http://www.nature.com/nclimate/journal/v4/n11/fig_tab/nclimate2385_T1.html (could make a binary chart for each state and run a regression)

```{r per capita state emission}
#https://www.eia.gov/environment/emissions/state/analysis/excel/table5.xlsx

if(file.exists("pcstate.xls")==FALSE){
  download.file("https://www.eia.gov/environment/emissions/state/analysis/excel/table5.xlsx",destfile="pcstate.xls")
pcstate<-read_excel("pcstate.xls")  
}else{
 pcstate<-read_excel("pcstate.xls")
}

View(pcstate)
```

```{r state gross co2 emission by year}
#https://www.eia.gov/environment/emissions/state/analysis/excel/table1.xlsx

if(file.exists("gross_state.xls")==FALSE){
  download.file("https://www.eia.gov/environment/emissions/state/analysis/excel/table1.xlsx",destfile="gross_state.xls")
gross_state<-read_excel("gross_state.xls",skip=4,col_names=FALSE)  
}else{
  gross_state<-read_excel("gross_state.xls",skip=4,col_names=FALSE)
}


```

```{r new variable state emission caps binary}
em_caps<-
#should i make an excel file on my computer and manually input emission caps as a binary variable?
```
take a look at the APIs and how we can use

can have multiple binary variables for the bivariate graphic NOA data (national oceanic something data online) and show different graphics of what's the average contaminants for states with this policy versus this policy etc

```{r univariate plot}
## Create a rank variable for income 
pd<-pd%>%mutate(percapinc_rank=rank(percapinc.2010))

## Plot by rank

gg<-ggplot(data=pd , aes(x=percapinc_rank,
                         y=percapinc.2010,
                         text=county))
##Add Axis Labels
gg<-gg+xlab("Rank")+ylab("Per Capita Income, 2010")

## Add Points
gg<-gg+geom_point(alpha=.5,size=.5)

gg

## Save for later
gg1<-gg

# Make Interactive plot
gg_p<-ggplotly(gg)

gg_p
```

```{r dif univariate plot (bar)}
gg<-ggplot(data,aes(x=Department,fill=Department))
gg<-gg+geom_bar()
gg
``` 

```{r histogram univariate}
gg<-ggplot(at,aes(x=DistanceFromHome))
gg<-gg+geom_histogram(binwidth = 1,fill="lightblue")
gg
``` 

```{r condtl_mean_single}
##Condtional Average across a single variable

## Create a variable for quartiles of college education
pd<-pd%>%mutate(coll_grad_level=ntile(coll_grad_pc,4))

table(pd$coll_grad_level)

pd<-pd%>%group_by(coll_grad_level)%>% ## Group by predictor
  ##Calculate mean at each level of predictor
  mutate(pred_income_college=mean(percapinc.2010))%>% 
  ## Ungroup
  ungroup()%>% 
  #Rank by prediction, with ties sorted randomly
  mutate(pred_income_college_rank=rank(pred_income_college,ties.method="random"))


gg<-ggplot(data=pd,aes(x=pred_income_college_rank,y=percapinc.2010,color="Actual"))
gg<-gg+geom_point(alpha=.5,size=.5)
gg<-gg+geom_point(aes(x=pred_income_college_rank,y=pred_income_college,color="Predicted:Conditional Mean, 1 var"))
gg<-gg+ scale_color_manual("Type",values=c("Predicted:Conditional Mean, 1 var"="red","Actual"="black"))
gg<-gg+theme(legend.position="bottom")
gg<-gg+xlab("Rank")+ylab("Per Capita Income, 2010")
gg

##Save for later
gg3<-gg

```

```{r condtl_mean_multiple}

##Condtional average across multiple variables

## Calculate quartiles of home value
pd<-pd%>%mutate(median_home_level=ntile(median_home_val,4))

table(pd$median_home_level)

##Make prediction: income by quartiles of education and home values

pd<-pd%>%group_by(coll_grad_level,median_home_level)%>% ## Grouping at multiple levels
  mutate(pred_income_college_home=mean(percapinc.2010))%>%
  ungroup()%>%
  mutate(pred_income_college_home_rank=rank(pred_income_college_home,ties.method="random"))

## Showing the various levels
gg<-ggplot(data=pd,aes(x=pred_income_college_home_rank,y=percapinc.2010,color="Actual"))
gg<-gg+geom_point(alpha=.5,size=.5)
gg<-gg+geom_point(aes(x=pred_income_college_home_rank,
                      y=pred_income_college_home,
                      color="Predicted:Conditional Mean, 2 vars")
)

gg<-gg+scale_color_manual("Type",values=c("Actual"="black",
                                  "Predicted:Conditional Mean, 2 vars"="orange"
))

gg<-gg+theme(legend.position="bottom")
gg<-gg+xlab("Rank")+ylab("Per Capita Income")

gg

## Save for later

gg4<-gg
```

```{r conditional mean plot w indep 2 variables}
## Summarize attrition by travel AND gender
at_sum<-at%>%
  group_by(BusinessTravel,Gender)%>%
  summarize(attr_avg=mean(attrit))

## Get the results
at_sum

## Plot it using a bar plot
gg<-ggplot(at_sum,aes(x=reorder(BusinessTravel,attr_avg),y=attr_avg,color=Gender))
gg<-gg+geom_bar(stat="identity",aes(fill=Gender),position="dodge")
gg
```

